---
title: "Improper Type Errors"
output: html_notebook
---

```{r setup, echo = FALSE}
library(dplyr)
library(readr)
library(tidyr)  
library(xtable)

options(dplyr.summarise.inform = FALSE)

all_path <- file.path("../all.csv")
all <- read_csv(
    all_path,
    show_col_types = FALSE
)
finished_early_path <- file.path("../locations/finished_early.csv")
finished_early <- read_csv(
    finished_early_path,
    show_col_types = FALSE
)
finished_late_path <- file.path("../locations/finished_late.csv")
finished_late <- read_csv(
    finished_late_path,
    show_col_types = FALSE
)
foreign_abis_path <- file.path("../locations/foreign_module_abis.csv")
foreign_abis <- read_csv(
    foreign_abis_path,
    show_col_types = FALSE
) %>% filter(abi == "C")
einfo_path <- file.path("../locations/error_info.csv")
einfo <- read_csv(
    einfo_path,
    show_col_types = FALSE
)
eloc_path <- file.path("../locations/error_locations.csv")
eloc <- read_csv(
    eloc_path,
    show_col_types = FALSE
)

discriminant_counts_path <- file.path("../locations/category_error_counts.csv")
discriminant_counts <- read_csv(
    discriminant_counts_path,
    show_col_types = FALSE
) %>% inner_join(einfo, by = c("err_id", "crate_name"))

discriminant_names_path <- file.path("../discriminants.csv")
discriminant_names <- read_csv(
    discriminant_names_path,
    show_col_types = FALSE
)

reasons_path <- file.path("../reasons.csv")
reasons <- read_csv(
    reasons_path,
    show_col_types = FALSE
)
lint_info_path <- file.path("../locations/lint_info.csv")
lint_info <- read_csv(
  lint_info_path,
  show_col_types = FALSE
)
```
```{r include = FALSE}
discriminant_counts %>%
  group_by(crate_name, category) %>%
  summarize(num_errs = sum(count)) %>% 
  group_by(category) %>% 
  summarize(
    num_crates = n(), 
    mean_errs = mean(num_errs), 
    max_errs = max(num_errs), 
    min_errs = min(num_errs), 
    stdev_errs = sd(num_errs)
  )

unique_errors <- einfo %>% 
  select(discriminant, err_text) %>% 
  unique %>% 
  nrow

unique_crates <- einfo %>% select(crate_name) %>% unique %>% nrow
```
We observed `r sum(discriminant_counts$count)` type errors in foreign bindings. There were `r unique_errors` unique errors which occurred for `r unique_crates` unique crates.

```{r echo = FALSE}
decl_disabled <- lint_info %>% filter(decl_disabled == TRUE) %>% select(crate_name) %>% unique
defn_disabled <- lint_info %>% filter(defn_disabled == TRUE) %>% select(crate_name) %>% unique
both_disabled <- lint_info %>% filter(defn_disabled == TRUE) %>% filter(decl_disabled == TRUE) %>% select(crate_name) %>% unique

g_decl_disabled <- decl_disabled %>% nrow # 40
g_defn_disabled <- defn_disabled %>% nrow # 11
g_both_disabled <- both_disabled %>% nrow # 10

partial <- discriminant_counts %>% select(crate_name, category, ignored)
partial <- partial %>% mutate(category = ifelse(as.character(category) %in% c('static_items','foreign_functions'), "decl", as.character(category)))
partial_defn <- partial %>% filter(category == 'rust_functions') %>% filter(ignored == TRUE) %>% select(crate_name) %>% unique # 12
partial_decl <- partial %>% filter(category == 'decl') %>% filter(ignored == TRUE) %>% select(crate_name) %>% unique #76
partial_both <- partial %>% filter(ignored == TRUE) %>% select(crate_name, category) %>% group_by(crate_name) %>% unique %>% summarize(n = n()) %>% filter(n == 2) %>% select(crate_name)

p_decl_disabled <- union(decl_disabled, partial_decl) %>% nrow # 80
p_defn_disabled <- union(defn_disabled, partial_defn) %>% nrow # 19
p_both_disabled <- union(both_disabled, partial_both) %>% nrow # 88
```
There were `r g_decl_disabled` crates that disabled the `improper_ctypes` lint. There were `r g_defn_disabled` crates that disabled the `improper_types_definitions` lint. There were `r g_both_disabled` crates that disabled both lints.

When we include crates that have lints disabled per-item, we have `r p_decl_disabled` crates that disabled `improper_ctypes` lint, we have `r p_defn_disabled` crates that disabled the `improper_types_definitions` lint, and `r p_both_disabled` crates that disabled both lints.
```{r echo = FALSE}

process_counts <- function(df) {
  df %>% group_by(crate_name, category) %>%
    summarize(num_errors = sum(count)) %>%
    group_by(category)
}

count_summary_all <- discriminant_counts %>%
  process_counts %>%
  summarize(
    n_crates_all = n(),
    errs_all = sum(num_errors),
    mean_errs_all = round(mean(num_errors), 1),
    stdev_errs_all = round(sd(num_errors), 1),
    max_errs_all = max(num_errors)
  )

unique_all <- discriminant_counts %>% select(crate_name, category, discriminant, reason, abi, err_text, ignored) %>%
  group_by(category) %>% distinct(crate_name, category, discriminant, reason, abi, err_text) %>% summarize(n_unique_all=n())
unique_ignored <- discriminant_counts %>% select(crate_name, category, discriminant, reason, abi, err_text, ignored) %>%
  group_by(category) %>% filter(ignored == TRUE) %>% distinct(crate_name, category, discriminant, reason, abi, err_text) %>% summarize(n_unique_ign=n())

count_summary_ign <-discriminant_counts %>% 
  filter(ignored == TRUE) %>%
  process_counts %>%
  summarize(
    n_crates_ign = n(),
    errs_ign = sum(num_errors),
    mean_errs_ign = round(mean(num_errors), 1),
    stdev_errs_ign = round(sd(num_errors), 1),
    max_errs_ign = max(num_errors)
  )

count_summary <- count_summary_all %>% inner_join(count_summary_ign, by=("category"))
count_summary$n_crates <- paste(count_summary_all$n_crates_all, "/", count_summary_ign$n_crates_ign)
count_summary$n_errs <- paste(count_summary_all$errs_all, "/", count_summary_ign$errs_ign)
count_summary$mean_errs <- paste(count_summary_all$mean_errs_all, "/", count_summary_ign$mean_errs_ign)
count_summary$stdev_errs <- paste(count_summary_all$stdev_errs_all, "/", count_summary_ign$stdev_errs_ign)
count_summary$max_errs <- paste(count_summary_all$max_errs_all, "/", count_summary_ign$max_errs_ign)

count_summary <- count_summary %>% inner_join(unique_all, by=("category"))
count_summary <- count_summary %>% inner_join(unique_ignored, by=("category"))
count_summary$n_unique <- paste(count_summary$n_unique_all, "/", count_summary$n_unique_ign)

count_summary <- count_summary %>% select(category, n_crates, n_unique, n_errs, mean_errs, stdev_errs, max_errs)
count_summary <- count_summary %>% mutate(category = ifelse(as.character(category) == "rust_functions", "Rust Functions", as.character(category)))
count_summary <- count_summary %>% mutate(category = ifelse(as.character(category) == "foreign_functions", "Foreign Functions", as.character(category)))
count_summary <- count_summary %>% mutate(category = ifelse(as.character(category) == "static_items", "Static Items", as.character(category)))

all_counts <- discriminant_counts %>% group_by(crate_name) %>%
    summarize(num_errors = sum(count))

all_counts_ign <- discriminant_counts %>% group_by(crate_name) %>%
  filter(ignored == TRUE) %>%
  summarize(num_errors = sum(count))

unique_ungrouped <- discriminant_counts %>% select(crate_name, category, discriminant, reason, abi, err_text, ignored) %>%
  distinct(crate_name, category, discriminant, reason, abi, err_text) %>% nrow

unique_ungrouped_ign <- discriminant_counts %>% select(crate_name, category, discriminant, reason, abi, err_text, ignored) %>%
  filter(ignored == TRUE) %>%
  distinct(crate_name, category, discriminant, reason, abi, err_text) %>% nrow

count_summary[nrow(count_summary) + 1,] <- list(
  'All', # Category
  paste(all_counts %>% nrow, '/', all_counts_ign %>% nrow), # Crates
  paste(unique_ungrouped, '/', unique_ungrouped_ign), # Crates
  paste(sum(all_counts$num_errors), '/', sum(all_counts_ign$num_errors)), # Unique Errors
  paste(round(mean(all_counts$num_errors), 1), '/', round(mean(all_counts_ign$num_errors), 1)), # Mean
  paste(round(sd(all_counts$num_errors), 1), '/', round(sd(all_counts_ign$num_errors), 1)), # St.Dev
  paste(max(all_counts$num_errors), '/', max(all_counts_ign$num_errors)) # Max
)
colnames(count_summary) <- c("Category", "# Crates", "# Unique Errors", "Total", "Mean", "St.Dev", "Max")

count_summary
```


```{r echo = FALSE}
named_counts <- discriminant_counts %>% 
  inner_join(discriminant_names, by=c("discriminant")) %>% filter(ignored == TRUE)

decls <- named_counts %>% filter(category %in% c("foreign_functions"))
defns <- named_counts %>% filter(category %in% c("rust_functions"))
decls_items <- named_counts %>% filter(category %in% c("static_items"))

discrim_counts <- discriminant_counts %>% 
  select(crate_name, category, discriminant, reason, abi, err_text, ignored) %>% 
  distinct(crate_name, category, discriminant, reason, abi, err_text) %>% 
  group_by(reason, discriminant) %>% 
  summarize(n_unique=n())

decls_items_counts <- decls_items %>%
  group_by(reason, discriminant) %>% 
  summarize(fi_count = sum(count))

decls_counts <- decls %>% 
  group_by(reason, discriminant) %>% 
  summarize(ff_count = sum(count))

defns_counts <- defns %>% 
  group_by(reason, discriminant) %>% 
  summarize(ef_count = sum(count))

sum(defns_counts$ef_count) + sum(decls_counts$ff_count) + sum(decls_items_counts$fi_count)


final <- decls_counts %>%
  full_join(decls_items_counts, by=c("discriminant","reason")) %>% 
  full_join(defns_counts, by=c("discriminant", "reason")) %>%
  full_join(discrim_counts, by=c("discriminant", "reason")) %>%
  inner_join(reasons, by=c("reason")) %>%
  inner_join(discriminant_names, by=c("discriminant")) %>%
  mutate(across(where(is.numeric), ~ replace_na(.x, 0))) %>%
  ungroup() %>%
  select(type_name, reason_name, n_unique, ff_count, fi_count, ef_count)

final$total_occurrences <- final$ff_count + final$fi_count + final$ef_count
filtered <- final %>% filter(total_occurrences > 0) %>% arrange(type_name)

filtered[nrow(filtered) + 1,] <- list(
  'All', 
  'All',
  sum(filtered$n_unique),
  sum(filtered$ff_count),
  sum(filtered$fi_count),
  sum(filtered$ef_count),
  sum(filtered$total_occurrences)
)
filtered <- filtered %>% mutate(reason_name = ifelse(as.character(type_name) == as.character(reason_name), "--", as.character(reason_name)))
colnames(filtered) <- c("Type", "Error", "Unique", "Foreign Functions", "Foreign Items", "Rust Functions", "Total")
filtered

```

```{r}

large <- function(x){
  paste0('\textbf{', x, '}')
}
addtorow <- list()
addtorow$pos <- list(0, 0)
addtorow$command <- c("& & & \\multicolumn{4}{c}{\\# Occurrences} \\\\\n",
"Category & \\# Crates & \\# Unique & Total & Mean & St.Dev. & Max \\\\\n")
xt_counts = xtable(
  count_summary, 
  type = "latex", 
  digits=c(0,0,0,0,0,1,1,0),

)

align(xt_counts) <- "llcc|cccc"
print(xt_counts, 
      include.rownames = FALSE,
      add.to.row = addtorow, 
      file = "../latex/err_counts.tex",
      include.colnames = FALSE,
      sanitize.colnames.function = large,
      floating=FALSE,
      latex.environments=NULL
)

addtorow <- list()
addtorow$pos <- list(0, 0)
addtorow$command <- c("& & & \\multicolumn{4}{c}{\\# Hidden Occurrences} \\\\\n",
"Error Category & Error & \\# Unique & Foreign Functions & Foreign Items & Rust Functions & Total \\\\\n")
xt_type_counts <- xtable(
  filtered, 
  type = "latex", 
  digits=c(0,0,0,0,0,0,0,0),
  floating=FALSE,
  latex.environments=NULL
)
align(xt_type_counts) <- "lllc|cccc"
print(xt_type_counts, 
      include.rownames = FALSE,
      add.to.row = addtorow, 
      file = "../latex/err_type_counts.tex",
      include.colnames = FALSE,
      sanitize.colnames.function = large,
      floating=FALSE,
      latex.environments=NULL
    )



```